---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![R build status](https://github.com/ThinkR-open/checkhelper/workflows/R-CMD-check/badge.svg)](https://github.com/ThinkR-open/checkhelper/actions)
[![Codecov test coverage](https://codecov.io/gh/ThinkR-open/checkhelper/branch/master/graph/badge.svg)](https://codecov.io/gh/ThinkR-open/checkhelper?branch=master)
<!-- badges: end -->
  
# checkhelper
A package to help you deal with `devtools::check()` outputs

Complete documentation in the {pkgdown} site: https://thinkr-open.github.io/checkhelper/

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


## Installation

You can install the last version of checkhelper from github with:

``` r
remotes::install_github("thinkr-open/checkhelper")
```

## Examples

- Check your current package under development and get all the globals missing: `no visible global variable` and `no visible global function`
- Detect exported functions with missing or empty `@return` tag

### Directly in your package in development

- Use `checkhelper::find_missing_tags()` on your package in development to find which functions are exported but missing `@export` roxygen2 tag. 
  - CRAN policy asks for every exported function to have a value (named `@export` when using {roxygen2}). 
  - This also checks that not exported functions dont have roxygen title, or have `@noRd`
- You can directly use `checkhelper::print_globals()` on your package instead of `devtools::check()`. This is a wrapper around `rcmdcheck::rcmdcheck()`. This will run the checks and directly list the potential "globalVariables" to add in a `globals.R` file.

```{r, eval=FALSE}
checkhelper::find_missing_tags()

checkhelper::print_globals(quiet = TRUE)
```

### Reproducible example with a fake package in tempdir

- Create a fake package with 
  - a function having global variables
  - a function with `@export` but no `@return`
  - a function with title but without `@export` and thus missing `@noRd`
  
```{r, eval=TRUE, results='hide'}
library(checkhelper)

# Create fake package ----
pkg_path <- tempfile(pattern = "pkg.")
dir.create(pkg_path)

# Create fake package
usethis::create_package(pkg_path, open = FALSE)

# Create function no visible global variables and missing documented functions
cat("
#' Function
#' @importFrom dplyr filter
#' @export
my_fun <- function() {
data %>%
filter(col == 3) %>%
mutate(new_col = 1) %>%
ggplot() +
  aes(x, y, colour = new_col) +
  geom_point()
}

#' Function not exported but with doc
my_not_exported_doc <- function() {
  message('Not exported but with title, should have @noRd')
}
", file = file.path(pkg_path, "R", "function.R"))

attachment::att_amend_desc(path = pkg_path)

# Files of the package
fs::dir_tree(pkg_path, recursive = TRUE)
```

- Find missing `@return` and find missing `@noRd` for not exported function with documentation

```{r}
find_missing_tags(pkg_path)
```


- Get global variables
```{r}
globals <- get_no_visible(pkg_path, quiet = TRUE)
globals
```

- Print globals to copy-paste
```{r, eval=TRUE}
print_globals(globals)
```

- Store the output of `print_globals()` in package using `usethis::use_r("globals")`. 
Note that you can also transform all these variables with `.data[[variable]]`


  Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
  By participating in this project you agree to abide by its terms.
