---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![R build status](https://github.com/ThinkR-open/checkhelper/workflows/R-CMD-check/badge.svg)](https://github.com/ThinkR-open/checkhelper/actions)
[![Codecov test coverage](https://codecov.io/gh/ThinkR-open/checkhelper/branch/master/graph/badge.svg)](https://codecov.io/gh/ThinkR-open/checkhelper?branch=master)
<!-- badges: end -->
  
# checkhelper
A package to help you deal with `devtools::check()` outputs

Complete documentation in the {pkgdown} site: https://thinkr-open.github.io/checkhelper/

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


## Installation

You can install the last version of checkhelper from github with:

``` r
remotes::install_github("thinkr-open/checkhelper")
```

## Examples

Check your current package under development and get all the globals missing: `no visible global variable` and `no visible global function`

- Create a fake package with a function having globalvariables
```{r, eval=TRUE, results='hide'}
library(checkhelper)

# Create fake package ----
tempdir <- tempdir()
# Create fake package
usethis::create_package(file.path(tempdir, "checkpackage"), open = FALSE)

# Create function no visible global variables and missing documented functions
cat("
#' Function
#' @importFrom dplyr filter
#' @export
my_fun <- function() {
data %>%
filter(col == 3) %>%
mutate(new_col = 1) %>%
ggplot() +
  aes(x, y, colour = new_col) +
  geom_point()
}
", file = file.path(tempdir, "checkpackage", "R", "function.R"))

path <- file.path(tempdir, "checkpackage")
attachment::att_to_description(path = path)
```

- Get global variables
```{r}
globals <- get_no_visible(path, quiet = TRUE)
globals
```

- Print globals to copy-paste
```{r, eval=TRUE}
print_globals(globals)
```

- Store in package using `usethis::use_r("globals")`


  Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
  By participating in this project you agree to abide by its terms.
