---
title: "flat_check_clean_userspace.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

# Check that the user space is clean after checks

`r lifecycle::badge("experimental")`

Have you faced a note on CRAN about non-standard things in the check directory ?

```
Check: for non-standard things in the check directory
Result: NOTE
    Found the following files/directories:
     ‘extrapackage’ 
```

Maybe you do not understand where these files came from.  
Then, you can run `check_clean_userspace()` in your package directory to detect every files that you created during the check.  
They could be issued from examples, tests or vignettes: `check_clean_userspace()` will tell you.


```{r function-check_clean_userspace}
#' check_clean_userspace
#' 
#' @description
#' `r lifecycle::badge("experimental")`
#'
#' @param pkg Path to package to check
#' @param check_output Path to directory where to store check results
#'
#' @return data.frame of files that are left after checks
#' @export
#'
#' @examples
check_clean_userspace <- function(pkg = ".", check_output = tempfile("dircheck")) {

  scratch_dir <- tempdir() #character(0)

  tmp_orig <- Sys.getenv("TMPDIR")
  Sys.setenv("TMPDIR" = scratch_dir)
  Sys.setenv("TMP" = scratch_dir)
  Sys.setenv("TEMP" = scratch_dir)

  on.exit(Sys.setenv("TMPDIR" = tmp_orig))

  if (!dir.exists(scratch_dir)) {dir.create(scratch_dir)}
  if (!dir.exists(check_output)) {dir.create(check_output)}

  all_files <- tibble(
    source = character(0),
    problem = character(0),
    where = character(0),
    file = character(0)
  )

  # browseURL(pkg)

  # Local snapshot
  local_tmpfile <- tempfile("local_shot")
  scratch_tmpfile <- tempfile("scratch_shot")
  local_shot <- utils::fileSnapshot(pkg, timestamp = local_tmpfile, md5sum = TRUE, recursive = TRUE, full.names = TRUE)
  scratch_shot <- utils::fileSnapshot(scratch_dir, timestamp = scratch_tmpfile, md5sum = TRUE, recursive = TRUE, full.names = TRUE)
  # home <- utils::fileSnapshot ("~", timestamp = tempfile("timestamp"), md5sum = TRUE)

    # Verify test do not leave files ----
    is.test <- list.files(file.path(pkg, "tests"))
    if (length(is.test) != 0) {
      cli::cli_rule("Unit tests")
      devtools::test(pkg = pkg, stop_on_failure = FALSE)
      all_files <- what_changed(local_shot, scratch_shot, source = "Unit tests", all_files, check_output = check_output)
    }

    # Update shots
    local_shot <- utils::fileSnapshot(pkg, timestamp = local_tmpfile, md5sum = TRUE, recursive = TRUE, full.names = TRUE)
    scratch_shot <- utils::fileSnapshot(scratch_dir, timestamp = scratch_tmpfile, md5sum = TRUE, recursive = TRUE, full.names = TRUE)

    # Verify examples do not leave files ----
    cli::cli_rule("Run examples")
    devtools::run_examples(pkg = pkg, run_dontrun = FALSE, run_dontrun = FALSE, fresh = FALSE, document = FALSE)
    all_files <- what_changed(local_shot, scratch_shot, source = "Run examples", all_files, check_output)
    if (any(all_files$source == "Run examples")) {
      warning("One of the 'Run examples' .R file was created to run examples. You should not bother about it")
    }

    # Update shots
    local_shot <- utils::fileSnapshot(pkg, timestamp = local_tmpfile, md5sum = TRUE, recursive = TRUE, full.names = TRUE)
    scratch_shot <- utils::fileSnapshot(scratch_dir, timestamp = scratch_tmpfile, md5sum = TRUE, recursive = TRUE, full.names = TRUE)

    # Verify Full check ----
    cli::cli_rule("Full check")
    rcmdcheck::rcmdcheck(path = pkg, check_output = check_output, args = c("--no-manual", "--as-cran"), quiet = TRUE)
    # browseURL(dircheck)
    all_files <- what_changed(local_shot, scratch_shot, source = "Full check", all_files, check_output)

    # Verify Full check - check dir ----
    pkgname <- read.dcf(file.path(pkg, "DESCRIPTION"))[,"Package"]

    the_dir <- list.files(file.path(check_output), pattern = paste0(pkgname, ".Rcheck"), full.names = TRUE)
    # Same tests, no new files in test during check
    still_files <- list.files(file.path(the_dir, "00_pkg_src", pkgname, "tests", "testthat"), full.names = TRUE)[
      !list.files(file.path(the_dir, "00_pkg_src", pkgname, "tests", "testthat")) %in%
        list.files(file.path(pkg, "tests", "testthat"))]

    if (length(still_files) != 0) {
      all_files <- all_files %>%
            rbind(
              tibble(
                source = "Tests in check dir",
                problem = "added",
                where = the_dir,
                file = still_files
              )
            )

      message("Some files were still in tests/testtthat dir after rcmdcheck: ", paste(still_files, collapse = ", "))
    }

    # Update shots
    local_shot <- utils::fileSnapshot(pkg, timestamp = local_tmpfile, md5sum = TRUE, recursive = TRUE, full.names = TRUE)
    scratch_shot <- utils::fileSnapshot(scratch_dir, timestamp = scratch_tmpfile, md5sum = TRUE, recursive = TRUE, full.names = TRUE)

    # After vignettes
    cli::cli_rule("Build Vignettes")
    the_v <- devtools::build_vignettes(pkg = pkg)
    if (!all(is.null(the_v))) {
      devtools::clean_vignettes(pkg = pkg)
      all_files <- what_changed(local_shot, scratch_shot, source = "Build Vignettes", all_files, check_output)
    }

  return(all_files)
}


#' @noRd
what_changed <- function(local_shot, scratch_shot, source, all_files, check_output) {

  # Ignore files issued from checks themselves
  file.no.problem <- paste0(
    normalizePath(check_output, winslash = "/"),
    "|[.]Rcheck/|",
    normalizePath(file.path(tempdir(), "callr-"), winslash = "/"),
    "|",
    normalizePath(file.path(tempdir(), "test.*[.](o|c|so)$"), winslash = "/"),
    "|",
    normalizePath(file.path(tempdir(), "foo[.]o$"), winslash = "/")
  )

  for (w.shot in c("local_shot", "scratch_shot")) {
    the_shot <- get(w.shot)
  # File changed
    all_local <- utils::changedFiles(the_shot, md5sum = TRUE, recursive = TRUE, full.names = TRUE)

    for (what in c("added", "deleted", "changed")) {
      if (length(all_local[[what]]) != 0) {
        # Not those in checkdir
        # grepl(normalizePath(file.path(tempdir(), "callr-"), winslash = "/"),
        #       normalizePath(all_local[[what]], winslash = "/"))
        all_local[[what]] <- normalizePath(all_local[[what]], winslash = "/")
        all_local[[what]] <- all_local[[what]][!grepl(file.no.problem, all_local[[what]])]
        if (length(all_local[[what]]) != 0) {
          all_files <- all_files %>%
            rbind(
              tibble(
                source = source,
                problem = what,
                where = the_shot$path,
                file = all_local[[what]]
              )
            )
          message("Some files were ", what," in ", the_shot$path, " after ", source, ": ",
                  paste(all_local[[what]], collapse = ", "))
        }
      }
    }
  }

  all_files
}

```

```{r examples-check_clean_userspace, eval=FALSE}
#' \dontrun{
all_files <- check_clean_userspace()
all_files
#' }
```

```{r tests-check_clean_userspace}
test_that("check_clean_userspace works", {

  if (!interactive()) {
    skip_on_os("windows")
    skip_on_os("mac")
  }

  path <- suppressWarnings(create_example_pkg())
  dir.create(file.path(path, "tests", "testthat"), recursive = TRUE)
  # Add a test that let file in the testthat dir
  cat(
    "cat('#in tests', file = 'in_test.R')", 
    file = file.path(path, "tests", "testthat", "test-in_test.R"))
  
  # Add an example that let file in tempdir
    cat(
    "#' Function
#' @return 1
#' @export
#' @examples
#' cat('#in example', file = tempfile('in_example'))
in_example <- function() {
1
}", 
    file = file.path(path, "R", "in_example.R"))
    
    # Add a vignette that let file in local ?
    
    suppressWarnings(attachment::att_amend_desc(path = path))
      
  check_output <- tempfile("check_output")
  scratch_dir <- tempfile("dirtmp")
  
  # debugonce(check_clean_userspace)
  expect_warning(
    expect_message(
      all_files <- check_clean_userspace(pkg = path, check_output = check_output),
      "Some files"),
    "One of the 'Run examples'")
  
  if (nrow(all_files) == 5) {
    # In some cases, the check updates the DESCRIPTION file as it run document()
    expect_equal(all_files$source, c("Unit tests", "Unit tests", "Run examples", "Run examples", 
                                     "Full check"))
    expect_equal(all_files$problem, c("added", "added", "added", "added", "added"))
    expect_equal(
      normalizePath(all_files$where, winslash = "/"),
      normalizePath(c(path, rep(tempdir(), 4)), winslash = "/"))
    expect_true(all(grepl("in_test[.]R", all_files$file[1:2])))
    expect_true(any(grepl("in_example", all_files$file[3:4]))) # One of the two
    expect_true(any(grepl("DESCRIPTION$", all_files$file[5])))
  } else if (nrow(all_files) == 6) {
    # In some cases, the check updates the DESCRIPTION file as it run document()
    expect_equal(all_files$source, c("Unit tests", "Unit tests", "Run examples", "Run examples", 
                                     "Full check", "Full check"))
    expect_equal(all_files$problem, c("added", "added", "added", "added", "added", "added"))
    expect_equal(
      normalizePath(all_files$where, winslash = "/"),
      normalizePath(c(path, rep(tempdir(), 5)), winslash = "/"))
    expect_true(all(grepl("in_test[.]R", all_files$file[1:2])))
    expect_true(any(grepl("in_example", all_files$file[3:4]))) # One of the two
    expect_true(any(grepl("DESCRIPTION$", all_files$file[5])))
    expect_true(any(grepl("symbols[.]rds$", all_files$file[6])))
  } else if (nrow(all_files) == 11) {
    expect_equal(all_files$source, c(
      "Unit tests", "Unit tests", "Run examples", "Run examples",
      "Full check", "Full check", "Full check", "Full check", "Full check",
      "Full check", "Full check"
    ))
    expect_equal(all_files$problem, c(
      "added", "added", "added", "added", "added", "added", "added",
      "added", "added", "added", "added"
    ))
    expect_equal(
      all_files$where,
      gsub(
        file.path("/private", c(path, rep(tempdir(), 10)), fsep = ""),
        pattern = "//",
        replacement = "/"
      )
    )
    expect_true(all(grepl("in_test[.]R", all_files$file[1:2])))
    expect_true(any(grepl("in_example", all_files$file[3:4]))) # One of the two
    expect_true(any(grepl("callr-", all_files$file[5:8])))
    expect_true(any(grepl("DESCRIPTION$", all_files$file[9])))
    expect_true(any(grepl("foo[.]o", all_files$file[10])))
    expect_true(any(grepl("symbols[.]rds", all_files$file[11])))
  } else {
    stop("Number of rows is not expected: ", all_files)
  }

  unlink(path, recursive = TRUE)
  unlink(check_output, recursive = TRUE)
  unlink(scratch_dir, recursive = TRUE)
})
```


```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(
  flat_file = "dev/flat_check_clean_userspace.Rmd", 
  vignette_name = "No files left after check", 
  check = FALSE, document = FALSE, open_vignette = FALSE
)
rstudioapi::navigateToFile("dev/0-dev_history.Rmd", line = 96)
```

