---
title: "flat_check_as_cran.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

# Check your package with CRAN settings

When you send your package on CRAN, there are multiple options set before running the checks. 
Here we use the CRAN settings and way of managing incoming packages used for Linux in this function `check_as_cran()`.  

Scripts and options used are directly issued from the GitHub mirror repository of the CRAN machines: <https://github.com/r-devel/r-dev-web/tree/master/CRAN/>.  
Although `check_as_cran()` should run on any OS, it will run CRAN parameters originally set up for Linux machines.  

In the `check_dir`, you will get the same outputs, in the same format as used by CRAN, for the pre-test of incoming packages.



```{r function-check_as_cran}
#' Check your package with real CRAN default environment variables and check strategy
#'
#' @description
#' `r lifecycle::badge("experimental")`
#' 
#' @param pkg pkg directory to check
#' @param check_dir Where to store check outputs
#' @param scratch Where to store temporary files (cleaned after)
#' @param Ncpus Number of Cpus
#' @param as_command Whether to run the check as Linux command line, instead of directly in R
#' @param clean_before Whether to delete the previous check_dir
#' @param open Whether to open the check dir at the end of the process
#'
#' @return An object containing errors, warnings, and notes.
#' 
#' @details
#' When you send your package on CRAN, there are multiple options set before running the checks. 
#' Here we use the CRAN settings and way of managing incoming packages used for Linux in this function `check_as_cran()`.  
#' 
#' Scripts and options used are directly issued from the GitHub mirror repository of the CRAN machines: <https://github.com/r-devel/r-dev-web/tree/master/CRAN/>.  
#' Although `check_as_cran()` should run on any OS, it will run CRAN parameters originally set up for Linux machines.  
#' 
#' In the `check_dir`, you will get the same outputs, in the same format as used by CRAN, for the pre-test of incoming packages.
#' 
#' @export
#'
#' @examples
check_as_cran <- function(pkg = ".", check_dir, scratch, Ncpus = 1, as_command = FALSE,
                          clean_before = TRUE,
                          open = FALSE) {
  pkg <- normalizePath(pkg)

  if (missing(check_dir)) {check_dir <- file.path(dirname(normalizePath(pkg)), "checkdir")}
  if (isTRUE(clean_before) | !dir.exists(check_dir)) {
    # if ( file.exists(check_dir)) {a <- try(file.remove(check_dir), silent = TRUE)}
    if (dir.exists(check_dir)) {
      unlink(check_dir, recursive = TRUE)
    }
    dir.create(check_dir)
  }

  if (missing(scratch)) {scratch <- file.path(dirname(normalizePath(pkg)), "scratch")}
  if (dir.exists(scratch)) {
    unlink(scratch, recursive = TRUE)
  } else {
    dir.create(scratch)
  }
  
    results <- the_check(
      pkg = pkg,
      check_dir =  check_dir,
      scratch = scratch,
      Ncpus = Ncpus,
      clean_before = clean_before
    )

  writeLines("\nDepends:")
  tools::summarize_check_packages_in_dir_depends(check_dir)
  writeLines("\nTimings:")
  tools::summarize_check_packages_in_dir_timings(check_dir)
  writeLines("\nResults:")
  tools::summarize_check_packages_in_dir_results(check_dir)
  writeLines("\nDetails:")
  tools::check_packages_in_dir_details(check_dir)
  message("\nSee all check outputs in: ", check_dir)
  
  if (isTRUE(open)) {
    utils::browseURL(check_dir)
  }
  return(results)
}
  

#' @noRd
the_check <- function(pkg = ".", check_dir, scratch, Ncpus = 1, as_command = FALSE, clean_before = TRUE) {

  pkgbuild::build(path = pkg, dest_path = check_dir)
  
  envfile <- system.file("cran/CRAN_incoming/check.Renviron", package = "checkhelper")
  # read.table does not read comment, which is good
  envfile_table <- utils::read.table(envfile, sep = "=")
  env_values <- as.list(t(envfile_table)[2,])
  names(env_values) <- as.list(t(envfile_table)[1,])
  
  lib_dir <- system.file("cran/lib", package = "checkhelper")
  
  withr::with_envvar(new = env_values, {
    if (!as_command) {
      Sys.setenv("TMPDIR" = scratch)
      check_unique(check_dir, lib_dir, scratch, Ncpus)
    } else {
      message("This may only run on Linux OS with sh")
      # Run as command line for Linux only
      lib_bin <- system.file("cran/bin", package = "checkhelper")
    }
  })
}
```

```{r examples-check_as_cran, eval=FALSE}
#' \dontrun{
# Directory to store the check outputs
check_dir <- tempfile("example")
# Check the current package
check_as_cran(check_dir = check_dir)
# Open directory with all outputs
utils::browseURL(check_dir)
#' }
```

```{r tests-check_as_cran}
test_that("check_as_cran works", {
  path <- suppressWarnings(create_pkg())
  check_dir <- tempfile("example")
  
  if (interactive()) {
    expect_error(check_as_cran(pkg = path, check_dir = check_dir), regexp = NA)
  } else {
    check_callr <- function(pkg, check_dir) {
      library(checkhelper)
      check_as_cran(pkg = pkg, check_dir = check_dir)
    }
    
    # need to run in another env. inside check in check
    expect_error(callr::r(check_callr, args = list(pkg = path, check_dir = check_dir)), regexp = NA)
  }
  # check_as_cran(pkg = "/mnt/Data/github/ThinkR-open/attachment")
  # Check that env is clean - only manually
  # Sys.getenv("_R_CHECK_AUTOCONF_")
  unlink(check_dir, recursive = TRUE)
})
```


```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_check_as_cran.Rmd", vignette_name = "Check with real CRAN settings", check = FALSE, document = FALSE)
rstudioapi::navigateToFile("dev/0-dev_history.Rmd", line = 96)
```

