---
title: "Auditing an R package you have just received"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Auditing an R package you have just received}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(checkhelper)
```

This vignette walks through the practical sequence of commands when a
colleague hands you an R package and asks "is this CRAN-ready?". The
goal is to surface every CRAN-blocking issue with the **smallest
possible number of `R CMD check` runs**.

## TL;DR — the audit script

```{r}
pkg <- "/path/to/the/package"

# 1. Run R CMD check ONCE and reuse it everywhere it's needed.
chk <- rcmdcheck::rcmdcheck(pkg, args = "--as-cran")

# 2. Static audits (no extra check needed).
audit_tags(pkg)            # exported funs without @return / internals without @noRd
audit_ascii(pkg)           # non-ASCII characters in R/, tests/, vignettes/, man/
audit_dataset_doc(pkg)     # datasets in data/ without a roxygen block

# 3. Audits that need the check output — pass `chk` to skip a 2nd run.
audit_globals(pkg, checks = chk)

# 4. Userspace audit (separate workflow: it runs tests + examples + check
#    with file-system snapshots between steps, can't reuse `chk`).
audit_userspace(pkg)

# 5. Apply the safe fixes.
fix_globals(pkg, checks = chk, write = TRUE)
fix_ascii(pkg, dry_run = TRUE)        # review first
fix_ascii(pkg, dry_run = FALSE)       # then apply
fix_dataset_doc(pkg)
```

## Why share the check object?

`audit_globals()` and `fix_globals()` parse the `notes` field of an
`rcmdcheck::rcmdcheck()` result to extract the
`no visible binding for global variable` and
`no visible global function definition` notes. By default each call
runs its own check, which is slow on a real package.

Both functions accept a `checks =` argument. When supplied, they skip
the `rcmdcheck()` call and parse the existing object. This lets you
run the check **once** and reuse the result for the whole audit.

```{r}
chk <- rcmdcheck::rcmdcheck(pkg, args = "--as-cran")

audit_globals(pkg, checks = chk)
fix_globals(pkg,   checks = chk, write = TRUE)
```

The other audits do not need a check at all:

| Audit                | Needs `R CMD check`?  | Notes                                  |
|----------------------|-----------------------|----------------------------------------|
| `audit_tags()`       | no                    | static via roxygen2                    |
| `audit_ascii()`      | no                    | AST via `getParseData()`               |
| `audit_dataset_doc()`| no                    | inspects `data/` and `R/`              |
| `audit_globals()`    | **yes** (reusable)    | accepts `checks =`                     |
| `audit_userspace()`  | yes (own pipeline)    | takes file-system snapshots, separate  |
| `audit_check()`      | yes                   | this **is** the check, with CRAN env   |

## Choosing between `rcmdcheck()` and `audit_check()`

- `rcmdcheck::rcmdcheck(pkg, args = "--as-cran")` — fast, returns the
  list shape that `audit_globals(checks = ...)` expects. Use it for the
  shared-check workflow above.
- `audit_check(pkg)` — runs the check with the full **CRAN incoming**
  environment (env vars, options, lib path mirrored from CRAN). Heavier
  and uses `tools::check_packages_in_dir()` under the hood; use it as a
  final gate before submission, not as the audit driver.

## Minimal end-to-end on a fake package

```{r}
pkg <- create_example_pkg()

chk <- rcmdcheck::rcmdcheck(pkg, args = "--as-cran")

audit_tags(pkg)
audit_ascii(pkg)
audit_dataset_doc(pkg)
audit_globals(pkg, checks = chk)

fix_globals(pkg, checks = chk, write = TRUE)
fix_ascii(pkg, dry_run = FALSE)
```

After applying the fixes, re-run the check (the package state has
changed, so a new `rcmdcheck()` is needed) and confirm 0 / 0 / 0.
