---
title: "Pre-submission gates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pre-submission gates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(checkhelper)
```

Once the dev-time audits described in
`vignette("auditing-an-r-package", package = "checkhelper")` are clean,
two heavier audits remain. They both have their **own pipeline** —
they cannot reuse the shared `chk` object — and you typically run
them last, just before `devtools::release()`.

| Audit                | What it does                                           | When to run            |
|----------------------|--------------------------------------------------------|------------------------|
| `audit_check()`      | `R CMD check` with the full CRAN incoming environment  | final gate             |
| `audit_userspace()`  | Detects files left behind by tests / examples / vignettes | before tagging a release |

## `audit_check()` — check with CRAN settings

`R CMD check` on your machine and `R CMD check` on CRAN's machines do
**not** see the same environment: CRAN sets a long list of env vars
and options before running the incoming-pretest. `audit_check()`
mirrors those settings (sourced from
<https://github.com/r-devel/r-dev-web/tree/master/CRAN/>) so a check
that is green locally has the highest chance of being green on CRAN.

It is heavier than a plain `rcmdcheck::rcmdcheck()` — use it as the
**final gate**, not as the audit driver during development.

```{r}
# Store the check outputs in a directory you can browse afterwards.
check_output <- tempfile("example")

# `Ncpus = 1` matches CRAN's incoming-pretest. Bump it on a multi-core
# host (e.g. `Ncpus = parallel::detectCores() - 1`) when you just need
# a fast pre-flight rather than a faithful CRAN reproduction.
audit_check(pkg = ".", check_output = check_output, Ncpus = 1)

# Open the directory with all outputs (logs, install, tests).
utils::browseURL(check_output)
```

The settings are originally tuned for the Linux pretest, so
`audit_check()` will run on any OS but the env it reproduces is the
Linux one.

## `audit_userspace()` — no files left after check

CRAN flags packages whose tests, examples or vignettes leak files
into the user space:

```
Check: for non-standard things in the check directory
Result: NOTE
    Found the following files/directories:
     'extrapackage'
```

`audit_userspace()` runs the package's tests, examples, vignettes and
full check, snapshots the file system between each step, and lists
every file that was created outside the check directory:

```{r}
leaks <- audit_userspace(pkg = ".")
leaks
```

Each row points to the source (`tests`, `examples`, `vignettes`,
`check`), the file, and where it landed. Use that to decide whether
the offending code should `withr::local_*()` its temp files,
`unlink()` on exit, or move under `tempdir()`.

This audit takes file-system snapshots that depend on the package's
real run, which is why it can't share the `rcmdcheck` object with
`audit_globals()`.
